RDF Constraint Types to Validate Metadata \\ on Highly-Complex Person-Level and Aggregated Data

### 2015-02-08 paper/paper7/paper7-appendix.tex
constraint type descriptions

\subsection{\underline{Data Model Consistency}}.
Is the data consistent with the intended semantics of the data model?
Such validation rules ensure the integrity of the data according to the data model.

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

\subsection{\underline{Data Model Consistency}}
Is the data consistent with the intended semantics of the data model?
Such validation rules ensure the integrity of the data according to the data model.

@@

In many cases, properties are used bi-directionally and then accessed in the inverse direction, e.g. parent $\equiv$ child$^{-}$. There should be a way to declare value type, cardinality etc of those inverse relations without having to declare a new property URI. 
The object property \emph{OP1} is an inverse\footnote{\emph{R-56-INVERSE-OBJECT-PROPERTIES}} of the object property \emph{OP2}. 
Thus, if an individual \emph{x} is connected by \emph{OP1} to an individual \emph{y}, then \emph{y} is also connected by \emph{OP2} to \emph{x}, and vice versa.

\begin{itemize}
	\item \textbf{{\em DISCO-C-INVERSE-OBJECT-PROPERTIES-01}}:
	\emph{disco:CategoryStatistics} resources are accessed from codes (\emph{skos:Concept}s) via \emph{disco:statisticsCategory}$^{-}$.
	\item \textbf{{\em DISCO-C-INVERSE-OBJECT-PROPERTIES-02}}:
	\emph{disco:SummaryStatistics} resources are accessed from \emph{disco:Variable}s via \emph{disco:statisticsVariable}$^{-}$.
	\item \textbf{{\em DISCO-C-INVERSE-OBJECT-PROPERTIES-03}}:
	\emph{disco:Variable}s are accessed from \emph{disco:Question}s via \emph{disco:question}$^{-}$.
\end{itemize}

@@

A role is symmetric if it is equivalent to its own inverse \cite{Kroetzsch2012}.
An object property symmetry axiom\footnote{\emph{R-61-SYMMETRIC-OBJECT-PROPERTIES}} states that the object property expression \emph{OPE} is symmetric - that is, if an individual \emph{x} is connected by \emph{OPE} to an individual \emph{y}, then \emph{y} is also connected by \emph{OPE} to \emph{x}.

@@

\begin{itemize}
	\item \textbf{{\em DISCO-C-DISJOINT-PROPERTIES-01:}} 
All \emph{Disco} properties (not having the same domain and range classes) are defined to be pairwise disjoint.
The following DL statements holds for each pair of \emph{Disco} properties (not having the same domain and range classes) ($variable \sqsubseteq \neg question$).
\end{itemize}

\begin{itemize}
	\item \textbf{{\em DATA-CUBE-C-DISJOINT-PROPERTIES-01:}} 
All \emph{Data Cube} properties (not having the same domain and range classes) are defined to be pairwise disjoint.
The following DL statements holds for each pair of \emph{Data Cube} properties (not having the same domain and range classes) ($dataSet \sqsubseteq \neg structure$).
\end{itemize}

@@

\emph{Inclusive or} is a logical connective joining two or more predicates that yields the logical value "true" when at least one of the predicates is true.
\emph{Context-Specific Inclusive OR of Properties}\footnote{\emph{R-202-CONTEXT-SPECIFIC-INCLUSIVE-OR-OF-PROPERTIES}} constraints specify that individuals are valid if they have at least one property relationship of one or multiple properties stated within a given context.
The context can be an application profile, a shape, or a class, i.e., the constraint applies for individuals of this specific class.

@@

At least one property group must match for individuals of a specific context. 
Context may be a class, a shape, or an application profile.

@@

Resource Shapes is a recursive language\footnote{\emph{R-222-RECURSIVE-QUERIES}} (the value shape of a Resource Shape is in turn another Resource Shape). 
There is no way to express that in SPARQL without hand-waving "and then you call the function again here" or "and then you embed this operation here" text.  
The embedding trick doesn't work
in the general case because SPARQL can't express recursive queries,
e.g. "test that this Issue is valid and all of the Issues that references, recursively".
Most SPARQL engines already have
functions that go beyond the official SPARQL 1.1 spec. 
The cost of that sounds manageable.

@@

An \emph{individual inequality axiom}\footnote{\emph{R-14-DISJOINT-INDIVIDUALS}} DifferentIndividuals( a$_1$ ... a$_n$ ) states that all of the individuals a$_i$, 1 $\leq$ i $\leq$ n, are different from each other; 
that is, no individuals a$_i$ and a$_j$ with i $\neq$ j can be derived to be equal. 
This axiom can be used to axiomatize the unique name assumption — the assumption that all different individual names denote different individuals.

@@

An \emph{equivalent object properties axiom}\footnote{\emph{R-4-EQUIVALENT-OBJECT-PROPERTIES}} \emph{EquivalentObjectProperties( OPE$_1$ ... OPE$_n$ )} states that all of the object property expressions \emph{OPE$_i$, 1 $\leq$ i $\leq$ n}, are semantically equivalent to each other. This axiom allows one to use each \emph{OPE$_i$} as a synonym for each \emph{OPE$_j$} — that is, in any expression in the ontology containing such an axiom, \emph{OPE$_i$} can be replaced with \emph{OPE$_j$} without affecting the meaning of the ontology. The axiom \emph{EquivalentObjectProperties( OPE$_1$ OPE$_2$ )} is equivalent to the following two axioms \emph{SubObjectPropertyOf( OPE$_1$ OPE$_2$ )} and \emph{SubObjectPropertyOf( OPE$_1$ OPE$_2$ )}.

An \emph{equivalent data properties axiom}\footnote{\emph{R-5-EQUIVALENT-DATA-PROPERTIES}} \emph{EquivalentDataProperties( DPE$_1$ ... DPE$_n$ )} states that all the data property expressions \emph{DPE$_i$, 1 $\leq$ i $\leq$ n}, are semantically equivalent to each other. This axiom allows one to use each \emph{DPE$_i$} as a synonym for each \emph{DPE$_j$} — that is, in any expression in the ontology containing such an axiom, DPE$_i$ can be replaced with \emph{DPE$_j$} without affecting the meaning of the ontology. The axiom \emph{EquivalentDataProperties( DPE$_1$ DPE$_2$ )} can be seen as a syntactic shortcut for the following axiom \emph{SubDataPropertyOf( DPE$_1$ DPE$_2$ )} and \emph{SubDataPropertyOf( DPE$_1$ DPE$_2$ )}.

\begin{itemize}
	\item \textbf{{\em DISCO-C-EQUIVALENT-PROPERTIES-01:}}
	Equivalent properties from different versions of \emph{Disco} can be marked as equivalent, e.g. \emph{disco:containsVariable} and \emph{disco:variable}. 
	As a consequence, the properties can be replaced by each other without affecting the meaning.
\end{itemize}

\begin{itemize}
	\item \textbf{{\em DATA-CUBE-C-EQUIVALENT-PROPERTIES-01:}}
	Equivalent properties from different versions of \emph{Data Cube} can be marked as equivalent. 
	As a consequence, the properties can be replaced by each other without affecting the meaning.
\end{itemize}

@@

\emph{{Property Assertions}}\footnote{\emph{R-96-PROPERTY-ASSERTIONS}}
and includes positive property assertions and negative property assertions.
A \emph{positive object property assertion ObjectPropertyAssertion( OPE a$_1$ a$_2$ )} states that the individual a$_1$ is connected by the object property expression \emph{OPE} to the individual a$_2$. 
A \emph{negative object property assertion NegativeObjectPropertyAssertion( OPE a$_1$ a$_2$ )} states that the individual a$_1$ is not connected by the object property expression \emph{OPE} to the individual a$_2$. 
A \emph{positive data property assertion DataPropertyAssertion( DPE a lt )} states that the individual a is connected by the data property expression DPE to the literal \emph{lt}. 
A \emph{negative data property assertion NegativeDataPropertyAssertion( DPE a lt )} states that the individual a is not connected by the data property expression \emph{DPE} to the literal \emph{lt}.

@@

Literals of given data properties within given contexts do not have to match given patterns\footnote{\emph{R-44-PATTERN-MATCHING-ON-RDF-LITERALS}}.

@@

Concept inclusions allow us to state that all mothers are female and that
all mothers are parents, but what we really mean is that mothers are exactly the female
parents. DLs support such statements by allowing us to form complex concepts such as
the \emph{intersection}\footnote{\ms{R-15-CONJUNCTION-OF-CLASS-EXPRESSIONS}, \ms{R-16-CONJUNCTION-OF-DATA-RANGES}} (also called \emph{conjunction})
which denotes the set of individuals that are both female and parents. A complex concept
can be used in axioms in exactly the same way as an atomic concept, e.g., in the
equivalence Mother $\equiv$ Female $\sqcap$ Parent .

@@

A \emph{complement class expression}\footnote{\emph{R-19-NEGATION-OF-CLASS-EXPRESSIONS}, \emph{R-20-NEGATION-OF-DATA-RANGES}} \emph{ObjectComplementOf( CE )} contains all individuals that are not instances of the class expression \emph{CE}.

@@

A \emph{minimum cardinality restriction}\footnote{{\em R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES}, {\em R-211-CARDINALITY-CONSTRAINTS}} contains all those individuals that are connected by a property to at least n different individuals/literals 
that are instances of a particular class or data range. If the class is missing, it is taken to be \emph{owl:Thing}. 
If the data range is missing, it is taken to be \emph{rdfs:Literal}.
$\leq n R. \top$ is the minimum unqualified cardinality restriction where $n \in \mathbb{N}$ (written $\leq  n R$ in short).
For unqualified cardinality restrictions, classes respective data ranges are not stated.

@@

A \emph{maximum cardinality restriction} contains all those individuals that are connected by a property to at most n different individuals/literals that are instances of a particular class or data range. If the class is missing, it is taken to be \emph{owl:Thing}. If the data range is not present, it is taken to be \emph{rdfs:Literal}.
Unqualified means that the class respective the data range is not stated. 
$\geq n R. \top$ is a \emph{maximum unqualified cardinality restriction}\footnote{\emph{R-82-MAXIMUM-UNQUALIFIED-CARDINALITY-ON-PROPERTIES}, \emph{R-211-CARDINALITY-CONSTRAINTS}} where $n \in \mathbb{N}$ (written $\geq  n R$ in short).

@@

A \emph{maximum cardinality restriction} contains all those individuals that are connected by a property to at most n different individuals/literals that are instances of a particular class or data range. If the class is missing, it is taken to be \emph{owl:Thing}. If the data range is not present, it is taken to be \emph{rdfs:Literal}.
Qualified means that the class respective the data range is stated. 
$\leq n R. C$ is a \emph{maximum qualified cardinality restriction}\footnote{\emph{R-76-MAXIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES}, \emph{R-211-CARDINALITY-CONSTRAINTS}} where $n \in \mathbb{N}$.

@@

\emph{Individual equality}\footnote{\emph{R-6-EQUIVALENT-INDIVIDUALS}} states that two different names are known to refer to the same individual \cite{Kroetzsch2012}.
